---
title: "Iris Dataset Case Study"
author: "Your Name"
date: "YYYY-MM-DD"
last_updated: "YYYY-MM-DD"
output:
  github_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

# Introduction  
This RMarkdown provides a minimal case study on the Iris dataset.  
We'll cover loading the data, exploratory visualization, building a simple classifier, and referencing external resources.

# Data Loading & Preview  
Below we load the Iris dataset, convert it to a data frame, and preview the first five rows.

```{r data-loading, echo=TRUE, message=FALSE}
# Load necessary libraries
library(datasets)
library(dplyr)

# Load the iris dataset
df <- as_tibble(iris)

# Display the first 5 rows
head(df, 5)
```

The code above:  
- Loads the built‑in `iris` dataset.  
- Converts it to a tibble for convenient printing.  
- Shows the first five rows of the data frame.

# Exploratory Visualization  
We create a simple scatter plot of two features to visualize class separation.

```{r exploratory-plot, echo=TRUE, message=FALSE}
library(ggplot2)

# Scatter plot of Sepal Length vs. Sepal Width
ggplot(df, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point() +
  labs(
    title = "Sepal Length vs. Sepal Width",
    x = "Sepal Length (cm)",
    y = "Sepal Width (cm)"
  )
```

The plot above shows the relationship between sepal length and width, colored by species.

# Minimal Working ML Example  
We'll split the data into training and testing sets, train a multinomial logistic regression classifier, and evaluate its performance.

```{r ml-example, echo=TRUE, message=FALSE}
library(caret)
library(nnet)

set.seed(42)
# Split into train/test
train_index <- createDataPartition(df$Species, p = 0.8, list = FALSE)
train <- df[train_index, ]
test  <- df[-train_index, ]

# Train a multinomial logistic regression model
model <- multinom(Species ~ ., data = train)

# Predict and evaluate
pred <- predict(model, test)
conf_mat <- confusionMatrix(pred, test$Species)

# Show results
print(conf_mat)
```

We chose a multinomial logistic regression (`nnet::multinom`) for its simplicity in handling multi‑class problems.  
The confusion matrix above summarizes the model's performance.

# How‑To Examples  
- [Go to Minimal Working ML Example](#minimal-working-ml-example)  
- To learn more about R’s built‑in datasets, see the [iris documentation](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/iris.html).

# External Resources  
Further reading: [Scikit‑learn Iris dataset documentation](https://scikit-learn.org/stable/modules/generated/sklearn.datasets.load_iris.html)
